The Talking Heads experiment

### 2015-04-29 chap10/chap10.tex
corrected some typos in Ch. 10-12

The Talking Heads experiments were obviously not an end point, but only the beginning. They confirmed the results that 
had been obtained with earlier theoretical models and computer simulations but did not push these models 
towards greater complexity. But subsequent work did expand the envelope of our understanding and modeling way  
beyond these boundaries, thanks to the work of many researchers who have since joined the research program. 
The expansion has happened along several dimensions: increased sophistication of the robots, a deeper theoretical 
understanding of semiotic dynamics, growth in the complexity of semantics and grammar, and new breakthroughs by 
studying the emergence and evolution of language strategies. An exhaustive survey of these exciting profound developments 
is beyond the scope of the present book. Instead, I will highlight here only some of the key language game experiments that 
used real physical robots and were direct variations or further extensions of the environments, game scripts
and strategies used in the Talking Heads experiment. This chapter focuses on experiments 
in the period before 2005, particularly work with the \textsc{aibo} robots and the first attempts towards the emergence 
of grammar.\footnote{An overview of these experiments is also given in: \cite{Steels:2005}.}
The next chapters focus on experiments after 2005.

@@

From 1996, even before the first Talking Heads experiment was started, experiments were already conducted in our 
laboratory, primarily by Paul Vogt, exploring language games on the cybernetic mobile robots that we were able to build ourselves
at the time. These robots were constructed from Lego bricks and 
used a basic processing board for linking directly simple sensors (touch sensors, infrared sensors) and actuators (left and right motors)
using an adaptive dynamical system (see \figref{f:plate3}). They were however too unreliable for long-term repeatable 
experimentation and the sensori-motor experiences were too restricted to hope for the development of interesting languages. The use 
of pan-tilt cameras, as in the Talking Heads experiment, was an attempt to 
have an experimental set-up at relatively low cost that was reliable and
used vision as the source of information about the world. Of course this came
at the price of less mobility and no true physical interaction with the real world. Nevertheless, many further fruitful experiments were 
done using the same set-up, particularly to explore in much greater depth the domain of color lexicons\footnote{Colour became a focal point of research because there is an extensive literature in cognitive science that has 
been gathering empirical data about color and its evolution and because it is relatively straightforward to do 
colour naming games. One of the main papers on our colour naming research is 
\cite{Steels:2005}.} and multi-word naming games.\footnote{Multi-word games were the focus of the Ph.D research by 
Joris Van Looveren, with a thesis defended in 2004.}
A Talking Heads simulator was also built by Paul Vogt in 
order to speed up such experiments and prepare the way for more advances with real robots \citep{Vogt:2003}.

@@

The robots start without any prior lexicon or ontology and have to come up with their own 
purely based on the feedback in the game and without any human intervention. 
They have been programmed to achieve autonomous locomotion and vision-based obstacle avoidance,
and maintain a real-time analog model of their immediate surroundings based on visual input. Using
this vision-based world model, the robots are able to detect and track
other robots as well as orange balls using standard image processing
algorithms (see \figref{fig:persp-rev-perception}). Furthermore, the
robots have been endowed with mechanisms to segment the flow of data
into distinct events and they have a short term memory in which they
can store a number of past events.

\begin{figure}[p]
  \centerline{\includegraphics[width=.90\textwidth]{chap10/figs/persp-rev-perception.pdf}}
\caption{\label{fig:persp-rev-perception}Chain from visual image (left) to an analog world model (middle) and a discrete event description
that is the topic of the language game (right). 
The top row shows this chain for robot A and the bottom row the chain for robot B, both looking at the same ball.}
\end{figure}

@@

The categories restricting slot-fillers are either themselves defined in terms of schemas
(for example, syn-cat-77 could be the restriction on a slot-filler in another schema), or they
are defined as rules that are applied in a forward-chaining fashion during matching. Two
examples of rules related to the above schema are:
\begin{verbatim}
  rule 101: ([WORD (W U)]) => ([MEMBER syn-cat-761])
  rule 99: ([WORD (W O)]) => ([MEMBER Syn-Cat-751])
\end{verbatim}
Semantic schemas describe the language-specific semantic structures underlying the
meanings of complete word groups. The closest linguistic correspondent to a semantic
schema is the notion of a case-frame. The constraints indicate how the total meaning
is constructed/decomposed into the meaning of the parts. During interpretation such
constraints therefore perform the same role as Montague style semantic interpretation
functions. The slots correspond to cases such as agent, patient, time, distance, or arguments
of semantic functions. The categories used to constrain what can fill a slot correspond
in linguistic terms to selection restrictions like animate, human, edible, future, etc. The
schema has also an associated category for the whole so that hierarchical combination is
possible. An example of a semantic schema is:
\begin{verbatim}
Schema-542
  SLOTS: (sem-slot-51 sem-slot-50)
  DESCRIPTION-SET: 
       ([sem-slot-50 sem-cat-751] [sem-slot-51 sem-cat-761])
  CONSTRAINTS: 
       ((CONJUNCTION (>> sem-slot-50) (>> sem-slot-51)))
  CATEGORY: sem-cat-77
  USE: 10
  SUCCESS: 3
\end{verbatim}

@@

But first we need a clearer idea on what case grammar is.
In the sentence \smplenquote{John gave Mary a book}, John plays the role of agent, the giver 
of the book, Mary that of recipient or 
beneficiary and the book is the \dblenquote{patient} or object of the action. Languages differ in terms of the strategies that they 
use to express these participant roles. Some of them, like German, have an elaborate system of morphological 
markers attached to constituents of the noun phrase. These markers signal that the 
phrase is nominative, accusative, dative, etc., and these cases then map in complex ways 
onto semantic roles such as agent and patient.
Other languages, like English, use a strategy based on word order and prepositions. For example, 
\smplenquote{Mary gave John a book} contains the same participants as \smplenquote{John gave Mary a book} but now Mary is the 
agent and John the recipient. Still other languages, like Japanese, use a strategy based on particles following 
the noun, as in the Japanese sentence \smplenquote{Tanaka-san wa Tokyo de o-to-san ni atta} 
meaning Tanaka met his father in Tokyo. Tanaka-san is the name of a person, oto-san means \smplenquote{father} and atta 
means \smplenquote{met}. The markers are \smplenquote{wa} for the topic, \smplenquote{de} for location and \smplenquote{ni} for another sense of location which can 
also function as possessive. The case grammar experiments focused on modeling the latter strategy.

@@

\begin{figure}[t]
  \centerline{\includegraphics[width=.40\textwidth]{chap10/figs/adam-eve.pdf}}
\centerline{\includegraphics[width=.70\textwidth]{chap10/figs/jack-give.pdf}}
\centerline{\includegraphics[width=.80\textwidth]{chap10/figs/push-jill.pdf}}
\caption{\label{fig:adam-eve}Top: Computer animation of the speaking and hearing agent used in public demonstrations 
of the case experiment. 
Middle: Example of a puppet scene, where Jill gives a block to Jack. It shows the source image (left) and a (partial) 
result of image processing (right). Bottom: Another example of a puppet scene where Jill pushes the block.}
\end{figure}
\clearpage
\subsubsection{Scaling up the vision system}
The vision system of the Talking Heads only dealt with static scenes. This had to be scaled up to
deal with dynamical scenes and event recognition. It was the focal work of 
Jean-Christophe Baillie at the Sony Computer Science Laboratory in Paris, who built a vision system 
called \textsc{peract} \is{\textsc{peract} vision system} inspired by an event-recognition developed by Jeffrey 
Siskind.\footnote{The vision system is described in the Ph.D thesis of J-C Baillie. A summary is found here: 
\cite{Steels:2003}. Inspiration for the event recognition system came 
primarily from the paper by \cite{Siskind:2000}.}
The vision system was decomposed into three subsystems with information flowing both in a bottom-up
(from image to world model) and top-down way (from world model and predictions to image): the subsystems achieved 
object identification, event identification, and derivation of qualitative descriptors.

@@

The detection and tracking of units at different hierarchical levels starts in a bottom-up manner from the images captured by the camera, and goes through various processing steps: figure/ground separation based on color perception, creation of a cellular occupancy grid, spatial region growing to identify spatial regions, and construction of spatio-temporal continuities based on 
color histograms of the objects (\figref{fig:PERACT-images}).

@@

The lexicon and grammar in \textsc{cogar} takes the form of a set of bi-directional rules. The rules have two poles 
which are typically the semantic and syntactic pole, although there could be rules that work only on the 
syntactic level and others that work on the semantic level. The rules are applied 
from right to left in production and from left to right in parsing. 
The rules can have variables at any position and in any of the two poles. In order to be applicable, 
one pole of the rule has to match, which means that the structure defined in the pole had to 
have correspondents within the target structure, possibly after binding the variables to concrete 
symbols in the coupled feature structure. If a complete match is 
found, the other pole is merged with the target, which means that all variables get instantiated and every 
element which is not yet in the target is added. The underlying computational model of \textsc{cogar}  
is therefore similar to a rule-based inference system, with an inference engine that cycles through the rules until 
no more rules can be applied. The variables are logic variables and hence the matching and merging is similar 
to the process of unification as used in PROLOG or other logic programming languages.

>>>>

The Talking Heads experiments were obviously not an end point, but only the beginning. They confirmed the results that 
had been obtained with earlier theoretical models and computer simulations but did not push these models 
towards greater complexity. But subsequent work did expand the envelope of our understanding and modelling way  
beyond these boundaries, thanks to the work of many researchers who have since joined the research programme. 
The expansion has happened along several dimensions: increased sophistication of the robots, a deeper theoretical 
understanding of semiotic dynamics, growth in the complexity of semantics and grammar, and new breakthroughs by 
studying the emergence and evolution of language strategies. An exhaustive survey of these exciting profound developments 
is beyond the scope of the present book. Instead, I will highlight here only some of the key language game experiments that 
used real physical robots and were direct variations or further extensions of the environments, game scripts
and strategies used in the Talking Heads experiment. This chapter focuses on experiments 
in the period before 2005, particularly work with the \textsc{aibo} robots and the first attempts towards the emergence 
of grammar.\footnote{An overview of these experiments is also given in: \cite{Steels:2005}.}
The next chapters focus on experiments after 2005.

@@

From 1996, even before the first Talking Heads experiment was started, experiments were already conducted in our 
laboratory, primarily by Paul Vogt, exploring language games on the cybernetic mobile robots that we were able to build ourselves
at the time. These robots were constructed from Lego bricks and 
used a basic processing board for linking directly simple sensors (touch sensors, infrared sensors) and actuators (left and right motors)
using an adaptive dynamical system (see \figref{f:plate3}). They were however too unreliable for long-term repeatable 
experimentation and the sensori-motor experiences were too restricted to hope for the development of interesting languages. The use 
of pan-tilt cameras, as in the Talking Heads experiment, was an attempt to 
have an experimental set-up at relatively low cost that was reliable and
used vision as the source of information about the world. Of course this came
at the price of less mobility and no true physical interaction with the real world. Nevertheless, many further fruitful experiments were 
done using the same set-up, particularly to explore in much greater depth the domain of colour lexicons\footnote{Colour became a focal point of research because there is an extensive literature in cognitive science that has 
been gathering empirical data about colour and its evolution and because it is relatively straightforward to do 
colour naming games. One of the main papers on our colour naming research is 
\cite{Steels:2005}.} and multi-word naming games.\footnote{Multi-word games were the focus of the Ph.D research by 
Joris Van Looveren, with a thesis defended in 2004.}
A Talking Heads simulator was also built by Paul Vogt in 
order to speed up such experiments and prepare the way for more advances with real robots \citep{Vogt:2003}.

@@

The robots start without any prior lexicon or ontology and have to come up with their own 
purely based on the feedback in the game and without any human intervention. 
They have been programmed to achieve autonomous locomotion and vision-based obstacle avoidance,
and maintain a real-time analogue model of their immediate surroundings based on visual input. Using
this vision-based world model, the robots are able to detect and track
other robots as well as orange balls using standard image processing
algorithms (see \figref{fig:persp-rev-perception}). Furthermore, the
robots have been endowed with mechanisms to segment the flow of data
into distinct events and they have a short term memory in which they
can store a number of past events.

\begin{figure}[p]
  \centerline{\includegraphics[width=.90\textwidth]{chap10/figs/persp-rev-perception.pdf}}
\caption{\label{fig:persp-rev-perception}Chain from visual image (left) to an analogue world model (middle) and a discrete event description
that is the topic of the language game (right). 
The top row shows this chain for robot A and the bottom row the chain for robot B, both looking at the same ball.}
\end{figure}

@@

The categories restricting slot-fillers are either themselves defined in terms of schemas
(for example, syn-cat-77 could be the restriction on a slot-filler in another schema), or they
are defined as rules that are applied in a forward-chaining fashion during matching. Two
examples of rules related to the above schema are:
\begin{verbatim}
  rule 101: ([WORD (W U)]) => ([MEMBER syn-cat-761])
  rule 99: ([WORD (W O)]) => ([MEMBER Syn-Cat-751])
\end{verbatim}
Semantic schemas describe the language-specific semantic structures underlying the
meanings of complete word groups. The closest linguistic correspondent to a semantic
schema is the notion of a case-frame. The constraints indicate how the total meaning
is constructed/decomposed into the meaning of the parts. During interpretation such
constraints therefore perform the same role as Montague style semantic interpretation
functions. The slots correspond to cases such as agent, patient, time, distance, or arguments
of semantic functions. The categories used to constrain what can fill a slot correspond
in linguistic terms to selection restrictions like animate, human, edible, future, etc. The
schema has also an associated category for the whole so that hierarchical combination is
possible. An example of a semantic schema is:
\clearpage
\begin{verbatim}
Schema-542
  SLOTS: (sem-slot-51 sem-slot-50)
  DESCRIPTION-SET: 
       ([sem-slot-50 sem-cat-751] [sem-slot-51 sem-cat-761])
  CONSTRAINTS: 
       ((CONJUNCTION (>> sem-slot-50) (>> sem-slot-51)))
  CATEGORY: sem-cat-77
  USE: 10
  SUCCESS: 3
\end{verbatim}

@@

But first we need a clearer idea on what case grammar is.
In the sentence \smplenquote{John gave Mary a book}, John plays the role of agent, the giver 
of the book, Mary that of recipient or 
beneficiary and the book is the \dblenquote{patient} or object of the action. Languages differ in terms of the strategies that they 
use to express these participant roles. Some of them, like German, have an elaborate system of morphological 
markers attached to constituents of the noun phrase. These markers signal that the 
phrase is nominative, accusative, dative, etc., and these cases then map in complex ways 
onto semantic roles such as agent and patient.
Other languages, like English, use a strategy based on word order and prepositions. For example, 
\smplenquote{Mary gave John a book} contains the same participants as \smplenquote{John gave Mary a book} but now Mary is the 
agent and John the recipient. Still other languages, like Japanese, use a strategy based on particles following 
the noun, as in the Japanese sentence \smplenquote{Tanaka-san wa Tokyo de o-to-san ni atta} 
meaning Tanaka met his father in Tokyo. Tanaka-san is the name of a person, oto-san means \smplenquote{father} and atta 
means \smplenquote{met}. The markers are \smplenquote{wa} for the topic, \smplenquote{de} for location and \smplenquote{ni} for another sense of location which can 
also function as possessive. The case grammar experiments focused on modelling the latter strategy.

@@

\begin{figure}[t]
  \centerline{\includegraphics[width=.40\textwidth]{chap10/figs/adam-eve.pdf}}
\centerline{\includegraphics[width=.70\textwidth]{chap10/figs/jack-give.pdf}}
\centerline{\includegraphics[width=.80\textwidth]{chap10/figs/push-jill.pdf}}
\caption{\label{fig:adam-eve}Top: Computer animation of the speaking and hearing agent used in public demonstrations 
of the case experiment. 
Middle: Example of a puppet scene, where Jill gives a block to Jack. It shows the source image (left) and a (partial) 
result of image processing (right). Bottom: Another example of a puppet scene where Jill pushes the block.}
\end{figure}

\subsubsection{Scaling up the vision system}
The vision system of the Talking Heads only dealt with static scenes. This had to be scaled up to
deal with dynamical scenes and event recognition. It was the focal work of 
Jean-Christophe Baillie at the Sony Computer Science Laboratory in Paris, who built a vision system 
called \textsc{peract} \is{\textsc{peract} vision system} inspired by an event-recognition developed by Jeffrey 
Siskind.\footnote{The vision system is described in the Ph.D thesis of J-C Baillie. A summary is found here: 
\cite{Steels:2003}. Inspiration for the event recognition system came 
primarily from the paper by \cite{Siskind:2000}.}
The vision system was decomposed into three subsystems with information flowing both in a bottom-up
(from image to world model) and top-down way (from world model and predictions to image): the subsystems achieved 
object identification, event identification, and derivation of qualitative descriptors.

@@

The detection and tracking of units at different hierarchical levels starts in a bottom-up manner from the images captured by the camera, and goes through various processing steps: figure/ground separation based on colour perception, creation of a cellular occupancy grid, spatial region growing to identify spatial regions, and construction of spatio-temporal continuities based on 
color histograms of the objects (\figref{fig:PERACT-images}).

@@

The lexicon and grammar in \textsc{cogar} takes the form of a set of bi-directional rules. The rules have two poles 
which are typically the semantic and syntactic pole, although there could be rules that work only on the 
syntactic level and others that work on the semantic level. The rules are applied 
from right to left in production and from left to right in parsing. 
The rules can have variables at any position and in any of the two poles. In order to be applicable, 
one pole of the rule has to match, which means that the structure defined in the pole had to 
have correspondents within the target structure, possibly after binding the variables to concrete 
symbols in the coupled feature structure. If a complete match is 
found, the other pole is merged with the target, which means that all variables get instantiated and every 
element which is not yet in the target is added. The underlying computational model of \textsc{cogar}  
is therefore similar to a rule-based inference system, with an inference engine that cycles through the rules until 
no more rules can be applied. The variables are logic variables and hence the matching and merging is similar 
to the process of unification as used in \textsc{prolog} or other logic programming languages.

### 2015-04-29 chap11/chap11.tex
corrected some typos in Ch. 10-12

In January 2007, Luc Steels, Martin Loetzsch, and Michael Spranger traveled to the Sony Computer Science Laboratory 
in Tokyo to do the first experiments with the \textsc{qrio} robots.  
We decided to redo the Talking Heads experiment, but focusing on proper names for individual objects, rather than the emergence 
of generic terms such as colours, shapes or positions.\enlargethispage{1\baselineskip}\footnote{The first paper on the Proper Naming Game is: \cite{Steels:2007}
It also discusses the concept of a semiotic network. 
A more recent paper is: \cite{Steels:2012b}. A more 
in-depth discussion of the experiment is included 
in the Ph.D thesis of Martin Loetzsch, defended in 2014 at Humboldt University.}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.85\textwidth]{chap10/figs/png-perception.pdf}}
\caption{\label{fig:png-perception}A robot is looking at consecutive scenes (from top to bottom) and tracking the objects in the scene. One object (labeled $O_{716}$)
is moved by a human experimenter.  We see from left to right: the original source image, results of region detection and 
colour segmentation, the (average) value for one dimension of one object ($O_{716}$), the segmented and labeled objects, 
and the world model, including the position of the other robot (indicated by a pointing arrow). }
\end{figure}
\clearpage
We call the feature vector for each object a {\itshape sensory experience}. \is{sensory experience}
It is represented using a spiking 
diagram, with a spike for each dimension and the value on that dimension marked (see 
\figref{fig:prototypes-identifiers-words}, left-most column). Real
world vision is difficult enough, but because the robots look at the
scene from different points of view, they often do not carve out the
same regions and they will certainly see different views of the
objects and locate them in different positions within their own
egocentric reference frame. So sensory experiences of the same scene
are always going to be different. This makes coordination and hence successful
communication very challenging.

@@

A semiotic network of an agent $a$ is defined as 
${\cal S}_a = O_a \times V_a \times I_a \times N_a$ where $O_a$ is the set of
sensory experiences of the agent grouped per scene, $V_a$ the set of
prototypical views maintained by $a$, $I_a$ the set of individuals
 known to $a$, and $N_a$ the set of names in $a$'s vocabulary
(see \figref{fig:prototypes-identifiers-words}
). Each link in the network is weighted (with a real number between 0.0 and 1.0). The
weight of the link between a sensory experience and a prototypical
view is based on nearest neighbor computation. The other weights are
stored in memory and reflect the confidence of the agent in the use of
that link based on past experience.

@@

The features used in the experiment focus on the upper torso of the robot and are based on a binary 
version of the original image. They rely on a standard pattern recognition technique, known as Normalised 
Centralised Moments (\cite{Hu:1962}). \is{Normalised Centralised Moments}
Moments are a global description of shape, capturing the statistical regularities
of its pixels for area, centre of mass, and orientation. Centralised moments are invariant to translation and 
normalised moments are invariant to translation and scale. Here we use seven moments, so that an image schema of 
the upper torso is captured in terms of a feature vector with seven data points, represented as a graph, 
although it could also be represented using the spike diagram used in the Proper Naming Game (\figref{vision}, right). 
These feature vectors are then classified using the same prototype-based approach as in the Proper Naming Game. A prototype of a posture 
consists of typical points for the seven moments, as well as a minimum and maximum deviation from 
each point. The best matching prototype is found by nearest neighbor computation. New prototypes are created 
when there is no prototype that matches distinctively with the sensory experience of a body posture and matching prototypes are
slightly adjusted to integrate new instances, so that the prototype progressively and adaptively reflects the body postures observed
in the world.

@@

To work out the procedural semantics hypothesis, we need to address the question what the primitives are of 
these semantic programs, and what computational formalism is most suited to combine them. Already in 2000, I reported
a representation system for computational semantics, called IRL.\footnote{
The earliest paper introducing IRL is: \cite{Steels:2000}. 
A more recent description of a new implementation is reported as: \cite{Spranger:2012}.}
IRL stands for Incremental Recruitment Language, for reasons that will become clear soon. \is{IRL}
IRL uses a constraint language approach. Constraint programming languages were pioneered in the  \is{Constraint Programming}
seventies by Borning, Steele, and others. It is a very active field with its own conferences, 
organisations and journals (see http://www.informatik.uni- trier.de/ ley/db/journals/constraints/) 
with wide applications in user interfaces, scheduling problems, design, games, etc.

The basic idea of a constraint language is that only dataflow is expressed explicitly, not control flow. 
There is a set of slots (variables) and a constraint attempts to reduce the set of possible 
fillers (bindings) for a slot. Constraints are combined in 
a constraint network, based on sharing slots. A solution is found when every slot in the network has unique bindings. 
Usually it requires looping several times through the constraints and often even a rewiring of the constraints to 
find a solution. A numerical example of a set of constraints on natural numbers could be:
\begin{verbatim}
  x = y + z with 2<y<5 and 0<x<5
\end{verbatim}
The constraints expressed here form a network because the same variables x and y (which are the slots of the 
network) appear in different constraints (the sum constraint and two instances of the <-constraint). 
Given this network, constraint propagation can compute easily that x = 4, y = 3, and z = 1.

\enlargethispage{2\baselineskip}
IRL formalises constraints using prefix notation in the following way. Slots are symbols preceded 
by question marks. Constraints are defined with a list with the first argument the name of the constraint and 
the remaining arguments the slots: 
\begin{verbatim}
 (<constraint> <arg1> ... <arg-n>)
\end{verbatim}
The first argument is called the target of the constraint because it is the most logical 'outcome' 
of applying a constraint, but as shown in the example above, data may flow in all directions. The numerical 
example given earlier translates into the following IRL program. 
\begin{verbatim}
(sum ?x ?y ?z)
(between ?y 2 5)
(between ?x 0 5)
\end{verbatim}

@@

So IRL is in fact a way to integrate the broad set of mechanisms used in 
pattern recognition (including neural networks) and AI (including machine learning). 
For example, an operation could implement a radial-basis-function network for doing categorisation and 
use the network to identify an object in a set. It could not only use the network but also adapt 
the weights in the network. Another operation could implement categorisation based 
on discrimination trees (as discussed in Chapter 4) and would not only do it, but also expand the discrimination trees 
when necessary. Other operations could implement set-theoretic operations, 
for the semantics of connectives like {\itshape and}, {\itshape or}, and {\itshape not}, by carrying out
{\itshape union}, {\itshape intersection} and {\itshape complement} operations over sets. There is no 
limit to the possible semantic operations, except that they have to terminate in finite time, possibly giving up when 
they cannot reach a valid result. IRL just provides a framework to define these operations and make them available 
to be recruited by the constraint network planner. IRL uses LISP as the host language, which means that all 
the computational machinery of LISP is available to implement semantic operations and associated learning algorithms.

@@

Interestingly, the best results in terms of communicative success
are obtained when the agents are allowed to evolve their own colour language, 
partly because there was no limit on the number of prototypes they could introduce. This was better 
than using the set of basic colour prototypes for English, simply because English would turn to more complex expressions
(such as \smplenquote{bluish green}) when the basic colour catgories are not effective. Moreover, 
agents still achieve better results compared to the English colour system when they do not need to stick 
dogmatically to it. Indeed this shows why alignment, which is widely observed
in human dialogues,\footnote{See: \cite{Garrod:1987} and \cite{Galantucci:2005}.}
is a good strategy for achieving more successful communication.

>>>>

In January 2007, Luc Steels, Martin Loetzsch, and Michael Spranger travelled to the Sony Computer Science Laboratory 
in Tokyo to do the first experiments with the \textsc{qrio} robots.  
We decided to redo the Talking Heads experiment, but focusing on proper names for individual objects, rather than the emergence 
of generic terms such as colours, shapes or positions.\enlargethispage{1\baselineskip}\footnote{The first paper on the Proper Naming Game is: \cite{Steels:2007}
It also discusses the concept of a semiotic network. 
A more recent paper is: \cite{Steels:2012b}. A more 
in-depth discussion of the experiment is included 
in the Ph.D thesis of Martin Loetzsch, defended in 2014 at Humboldt University.}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.85\textwidth]{chap10/figs/png-perception.pdf}}
\caption{\label{fig:png-perception}A robot is looking at consecutive scenes (from top to bottom) and tracking the objects in the scene. One object (labelled $O_{716}$)
is moved by a human experimenter.  We see from left to right: the original source image, results of region detection and 
colour segmentation, the (average) value for one dimension of one object ($O_{716}$), the segmented and labelled objects, 
and the world model, including the position of the other robot (indicated by a pointing arrow). }
\end{figure}
\clearpage
We call the feature vector for each object a {\itshape sensory experience}. \is{sensory experience}
It is represented using a spiking 
diagram, with a spike for each dimension and the value on that dimension marked (see 
\figref{fig:prototypes-identifiers-words}, left-most column). Real
world vision is difficult enough, but because the robots look at the
scene from different points of view, they often do not carve out the
same regions and they will certainly see different views of the
objects and locate them in different positions within their own
egocentric reference frame. So sensory experiences of the same scene
are always going to be different. This makes coordination and hence successful
communication very challenging.

@@

A semiotic network of an agent $a$ is defined as 
${\cal S}_a = O_a \times V_a \times I_a \times N_a$ where $O_a$ is the set of
sensory experiences of the agent grouped per scene, $V_a$ the set of
prototypical views maintained by $a$, $I_a$ the set of individuals
 known to $a$, and $N_a$ the set of names in $a$'s vocabulary
(see \figref{fig:prototypes-identifiers-words}
). Each link in the network is weighted (with a real number between 0.0 and 1.0). The
weight of the link between a sensory experience and a prototypical
view is based on nearest neighbour computation. The other weights are
stored in memory and reflect the confidence of the agent in the use of
that link based on past experience.

@@

The features used in the experiment focus on the upper torso of the robot and are based on a binary 
version of the original image. They rely on a standard pattern recognition technique, known as Normalised 
Centralised Moments (\cite{Hu:1962}). \is{Normalised Centralised Moments}
Moments are a global description of shape, capturing the statistical regularities
of its pixels for area, centre of mass, and orientation. Centralised moments are invariant to translation and 
normalised moments are invariant to translation and scale. Here we use seven moments, so that an image schema of 
the upper torso is captured in terms of a feature vector with seven data points, represented as a graph, 
although it could also be represented using the spike diagram used in the Proper Naming Game (\figref{vision}, right). 
These feature vectors are then classified using the same prototype-based approach as in the Proper Naming Game. A prototype of a posture 
consists of typical points for the seven moments, as well as a minimum and maximum deviation from 
each point. The best matching prototype is found by nearest neighbour computation. New prototypes are created 
when there is no prototype that matches distinctively with the sensory experience of a body posture and matching prototypes are
slightly adjusted to integrate new instances, so that the prototype progressively and adaptively reflects the body postures observed
in the world.

@@

To work out the procedural semantics hypothesis, we need to address the question what the primitives are of 
these semantic programs, and what computational formalism is most suited to combine them. Already in 2000, I reported
a representation system for computational semantics, called IRL.\footnote{
The earliest paper introducing IRL is: \cite{Steels:2000}. 
A more recent description of a new implementation is reported as: \cite{Spranger:2012}.}
IRL stands for Incremental Recruitment Language, for reasons that will become clear soon. \is{IRL}
IRL uses a constraint language approach. Constraint programming languages were pioneered in the  \is{Constraint Programming}
seventies by Borning, Steele, and others. It is a very active field with its own conferences, 
organisations and journals (see \url{http://www.informatik.uni-trier.de/ley/db/journals/constraints/}) 
with wide applications in user interfaces, scheduling problems, design, games, etc.

The basic idea of a constraint language is that only data flow is expressed explicitly, not control flow. 
There is a set of slots (variables) and a constraint attempts to reduce the set of possible 
fillers (bindings) for a slot. Constraints are combined in 
a constraint network, based on sharing slots. A solution is found when every slot in the network has unique bindings. 
Usually it requires looping several times through the constraints and often even a rewiring of the constraints to 
find a solution. A numerical example of a set of constraints on natural numbers could be:
\begin{verbatim}
  x = y + z with 2<y<5 and 0<x<5
\end{verbatim}
The constraints expressed here form a network because the same variables x and y (which are the slots of the 
network) appear in different constraints (the sum constraint and two instances of the <-constraint). 
Given this network, constraint propagation can compute easily that x = 4, y = 3, and z = 1.

\enlargethispage{2\baselineskip}
IRL formalises constraints using prefix notation in the following way. Slots are symbols preceded 
by question marks. Constraints are defined with a list with the first argument the name of the constraint and 
the remaining arguments the slots: 
\begin{verbatim}
 (<constraint> <arg1> ... <arg-n>)
\end{verbatim}
The first argument is called the target of the constraint because it is the most logical ``outcome" 
of applying a constraint, but as shown in the example above, data may flow in all directions. The numerical 
example given earlier translates into the following IRL program. 
\begin{verbatim}
(sum ?x ?y ?z)
(between ?y 2 5)
(between ?x 0 5)
\end{verbatim}

@@

So IRL is in fact a way to integrate the broad set of mechanisms used in 
pattern recognition (including neural networks) and AI (including machine learning). 
For example, an operation could implement a radial-basis-function network for doing categorisation and 
use the network to identify an object in a set. It could not only use the network but also adapt 
the weights in the network. Another operation could implement categorisation based 
on discrimination trees (as discussed in \chapref{chap:4}) and would not only do it, but also expand the discrimination trees 
when necessary. Other operations could implement set-theoretic operations, 
for the semantics of connectives like {\itshape and}, {\itshape or}, and {\itshape not}, by carrying out
{\itshape union}, {\itshape intersection} and {\itshape complement} operations over sets. There is no 
limit to the possible semantic operations, except that they have to terminate in finite time, possibly giving up when 
they cannot reach a valid result. IRL just provides a framework to define these operations and make them available 
to be recruited by the constraint network planner. IRL uses LISP as the host language, which means that all 
the computational machinery of LISP is available to implement semantic operations and associated learning algorithms.

@@

Interestingly, the best results in terms of communicative success
are obtained when the agents are allowed to evolve their own colour language, 
partly because there was no limit on the number of prototypes they could introduce. This was better 
than using the set of basic colour prototypes for English, simply because English would turn to more complex expressions
(such as \smplenquote{bluish green}) when the basic colour categories are not effective. Moreover, 
agents still achieve better results compared to the English colour system when they do not need to stick 
dogmatically to it. Indeed this shows why alignment, which is widely observed
in human dialogues,\footnote{See: \cite{Garrod:1987} and \cite{Galantucci:2005}.}
is a good strategy for achieving more successful communication.
